---
title: "Workshop 7: Pivoting Demo"
output:
  prettydoc::html_pretty:
    theme: architect
    toc: true
    number_sections: true
author: "Author's name here"
date: "2022-11-22"
editor_options: 
  chunk_output_type: inline
  markdown: 
    wrap: 72
---

```{r include = FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Load packages 
if(!require(pacman)) install.packages("pacman")
pacman::p_load(tidyverse, here)

```

# Pivoting Next Level

## Wide to Long

Consider the following [data set from Zambia](https://zenodo.org/record/4571669) about enteropathogens and their biomarkers.

```{r}
enteropathogens_zambia_wide<- read_csv(here("data/enteropathogens_zambia_wide.csv"))

enteropathogens_zambia_wide
```

This data frame has the following columns:

-   LPS_1 and LPS_2: lipopolysaccharide levels, measured by Pyrochrome LAL, in EU/mL

-   LBP_1 and LBP_2: LPS binding protein levels, in pg/mL

-   IFABP_1 and IFAPB_2: intestinal-type fatty acid binding protein levels, in pg/mL

Pivot the dataset so that it resembles the following structure

```{r}
enteropathogens_zambia_long <- 
  enteropathogens_zambia_wide %>% 
  pivot_longer(!ID, 
               names_to = c(".value", "sample_count"), 
               names_sep = "_")

enteropathogens_zambia_long
```

## Long to Wide

```{r}
enteropathogens_zambia_wide <-
  enteropathogens_zambia_long %>%
  pivot_wider(
    names_from = sample_count,
    values_from = c(LPS, LBP, IFABP)
  )

enteropathogens_zambia_wide
```
