---
title: "Honors Assignment: Creating interactive ggplots"
output:
  prettydoc::html_pretty:
    theme: architect
    toc: true
    number_sections: true
author: "Author's name here"
date: "2023-01-01"
editor_options: 
  chunk_output_type: console
---

```{r include = FALSE}
library(knitr)
```

# Intro

Welcome!

For this assignment, you will need to create a ggplot graph and add interactivity to it using the **{plotly}** package.  https://plotly.com/ggplot2/.

The final due date for the assignment is Tuesday, January 31th at 23:59 PM UTC+2.


# Getting started

Let's begin by loading the packages you will need to use:

```{r}
# Load packages 
if(!require(pacman)) install.packages("pacman")
pacman::p_load(plotly, # for adding interactivity!
               tidyverse, # for ggplot2 and dplyr
               htmlwidgets, # for saving interactive graph
               gapminder # for data
               )
```
    
We have not yet used {plotly} in this course. This code will install the package if you don't have it already.

Next, we will load in and prepare data.

In this lesson, we will be using the `gapminder` dataset, but only the data from 2007.

```{r}
# Import data frame from gapminder package
gapminder07 <- gapminder::gapminder %>%
  # filter data frame to only include data from 2007
  filter(year == 2007) %>%
  # remove the year column
  select(-year) %>%
  # rename columns to make them easier to understand
  rename(life_expectancy = lifeExp,
         population = pop,
         gdp_per_capita = gdpPercap)
```


```{r}
# View dataset
head(gapminder07, 10)
```


# Intro to plotly

The **plotly** package is used to create interactive graphs in R and Python. It can used to make a variety of plot types, just like ggplot. 

An interactive chart allows you to change what the chart shows by taking actions: pan and zoom, hover over a geom to get additional information, choose which variables to display, and more.

Let's say we want to view the relationship between life expectancy and GPD per capita in `gapminder07` using a interactive scatter plot. 


You *could* learn plotly syntax use the `plot_ly()` function to do this in one step (but we will not be using this method):

```{r}
# Interactive scatter plot using plot_ly!
plot_ly(data = gapminder07,
        type = "scatter",
        x = ~gdp_per_capita, 
        y = ~life_expectancy)
```

The interactive plot will appear in your Viewer pane, and you can hover over the points to see the raw data values for x and y. 

But we don't need to do this and learn all that syntax, instead we will use the `ggplotly()` function that makes this super easy!

**`ggplotly()`** is a function converts ggplot objects into interactive plots in two simple steps:

1.  Create a named ggplot2 plot using the **`ggplot()`** function and the assignment operator (`<-`).

2.  Use the **`ggplotly()`** function to convert the ggplot2 plot into an interactive plotly plot: **`ggplotly(PLOT_NAME)`**

**Step 1**

```{r}
# First create ggplot object
base_ggplot <- 
  ggplot(data = gapminder07,
         mapping = aes(x = gdp_per_capita,
                       y = life_expectancy)) +
  geom_point()
```

**Step 2**

```{r}
# Transform plot with ggplotly()
ggplotly(base_ggplot)
```

## Labels

We can add country labels to hover text using the labels aesthetic inside `aes()`.

```{r}

# Create a ggplot2 plot with data labels for the country variable
gapminder_plot <- ggplot(
  data = gapminder07,
  mapping = aes(
    x = gdp_per_capita,
    y = life_expectancy,
    label = country)) +
  geom_point()

# Convert the ggplot2 plot to an interactive plotly plot

ggplotly(gapminder_plot)
```


# Requirements

- use at least 3 aesthetic mappings 
- at least one fixed aesthetic
- edit scale or position
- use the `labs()` function


# Submission: Upload zipped folder

Once you have finished the tasks above, you should **create a zipped file** of your project folder, containing all the work you have done, and **upload it** on the assignment page [LINK].

The final due date for the assignment is Tuesday, January 31th at 23:59 PM UTC+2.