---
title: "Workshop 10: Histograms, positions and facets Exercise"
output:
  prettydoc::html_pretty:
    theme: architect
    toc: true
    number_sections: true
author: "Your name here"
date: "2022-12-13"
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r include = FALSE}
# Load packages 
if(!require(pacman)) install.packages("pacman")
pacman::p_load(tidyverse, gapminder, here)
```

Welcome to your last ggplot2 live workshop!

In this exercise you will be using a subset of the `gapminder`
dataframe. Create a subset called `gap_small`, which only contains data
from 1952 and 2007.

```{r}
gap_small <- "YOUR CODE HERE"
```

First, plot a simple histogram showing the distribution of life
expectancy (lifeExp) in your dataframe.

```{r}
ggplot()
```

Set `binwidth = 5` to give each bar a 5-year range (and get rid of the
persistent warning message).

```{r}
ggplot()
```

Use the `fill` argument to create a stacked histogram with two fill
colors, one for each year.

```{r}
ggplot()
```

Hint: year is treated as numerical, so `ggplot()` will try to map it as
a continuous variable. To get two distinct colors, you will need to tell
R to treat it as a factor.

Change the position argument to overlap the two distributions. Then, add
a degree of transparency so that you can see where the bars overlap.

```{r}
ggplot()
```

Create small multiples of the plot above using `facet_wrap()`, with one
panel for each continent.

```{r}
ggplot()
```

Next, use `facet_grid()` to further subdivide your plots by `year`. You
should have one column for each continent, and one row for each year.

```{r}
ggplot()
```

There are only two countries in the continent "Oceania". Remove these
from your dataframe (or if you want an extra challenge, use `mutate()`
and `case_when()` to change the continent name from "Oceania" to
"Asia").

```{r}
gap_smaller <- gap_small %>% 
  mutate(continent = case_when(continent == "Oceania" ~ "Asia",
                          TRUE ~ as.character(continent)))
```

Recreate your previous plot with the new dataframe. This time, map
continent to fill color.

```{r}

ggplot()

```

Lastly, improve the labels of the axes and color legend on this plot.

```{r}

```

# Submission: Upload Rmd and HTML

The final due date for this exercise is Wednesday, December 14th at
23:59 PM UTC+2.

Once you have finished the tasks above, you should **knit this Rmd into
an HTML** and **upload both files** on the assignment page in a ZIP
folder.
