---
title: "Workshop 6: Grouping, summarizing and plotting"
output:
  prettydoc::html_pretty:
    theme: architect
    toc: true
    number_sections: true
author: "Zachary Anderson"
date: "2022-11-22"
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```



# Intro

Welcome!

For this workshop, we will be cleaning a dataset. It is a hands-on
approach to using the `select()`, `filter()`, `mutate()`, `case_when()`, and `summarize()`.

The assignment should be submitted individually, but you are encouraged
to brainstorm with partners.

The final due date for the assignment is Tuesday, November 22th at 23:59
PM UTC+2.

# Get the assignment repo

To get started, you should download and look through the assignment
folder.

1.  First download the repo to your local computer
    [here](https://minhaskamal.github.io/DownGit/#/home?url=https://github.com/the-graph-courses/rbp_cohort_0_materials/tree/main/week_06){target="_blank"}.

    *You should ideally work on your local computer, but if you would
    rather work on RStudio Cloud, you can upload the zip file to RStudio
    Cloud through the Files pane. Consult one of the instructors for
    guidance on this.*

2.  Unzip/Extract the downloaded folder.

    *If you are on macOS, you can simply double-click on a file to unzip
    it.*

    *If you are on Windows and are not sure how to "unzip" a file, see
    [this image](https://imgur.com/a/6e5pT7k){target="_blank"}. You need
    to right-click on the file and then select "extract all".*

3.  Once done, click on the RStudio Project file in the unzipped folder
    to open the project in RStudio.

4.  In RStudio, navigate to the Files tab and open the "rmd" folder. The
    instructions for your exercise are outlined there (these are the
    same instructions you see here).

5.  Open the "data" folder and observe its components. You will work
    with the "india_tuberculosis.csv" file. (You can also open the
    "00_info_about_the_dataset" file to learn more about this dataset.)




```{r}
library(esquisse)
library(here)
library(tidyverse)

data <- read.csv(here("data/india_tuberculosis.csv"))

data <- data %>%
  mutate(bmi_category = case_when(bmi>30 ~ "Obese"),
         total_cost = 1000) %>%
  group_by(Sex, bmi_category) %>%
  summarize(mean_total_cost = mean(total_cost)) %>%
  filter(bmi_category=="Obese")%>%
  select(Sex, mean_total_cost)



# esquisser(data)

```


# Load packages and data

Now that you understand the structure of the repo, you can load in and
clean your dataset.

In the code section below, **load in the needed packages**.

```{r}
if(!require(pacman)) install.packages("pacman")
pacman::p_load(here, janitor, reactable, readr, tidyverse)
```

Now, **read the dataset into R**. The data frame you import should have
880 rows and 22 columns. Remember to use the `here()` function to allow
your Rmd to use project-relative paths.

**Pro-tip:** column names aren't standardized this time. I would recommend using `janitor::clean_names()` (the function `clean_names()` from the `janitor` package). A doubt about how to write it? Go back to week 5, the mutate example. 

```{r}
india_TB <- read.csv(here("data/india_tuberculosis.csv"))

tuberculosis_data_cleaned <- clean_names(india_TB)
```

# Create new variables

## **Step 1:** Encode an `age_group` variable

The age variable represents the age of participants in years. For further manipulations, we want to create an `age_group` variable with the following categories:

-   \<10

-   10-17

-   18-29

-   30-49

-   50-79

-   80+

```{r}
# Making the categories; note that the "clean_names()" function above made "Age" lowercase

tuberculosis_data_cleaned <- mutate(tuberculosis_data_cleaned,
                    age_group = case_when(age <10 ~ "Prepubescent",
                    age >=10 & age <=17 ~ "Teen",
                    age >=18 & age <=29 ~ "Young adult",
                    age >=30 & age <=49 ~ "Adult",
                    age >=50 & age <=79 ~ "Mature adult",
                    age >=80 ~ "Senior"
                    ))

```

## **Step 2:** Create `bmi` variableS

### Part A: Calculate the BMI 

You have at your disposal the weight (in kg) and height (in cm) of your participants. Calculate the BMI of your participants.

(Careful ! Check your units for weight and height: are your variables in the right unit or do you need to convert them to another unit using `mutate()`? )

```{r}
# again note the names changed by "clean_names()"
tuberculosis_data_cleaned <- mutate(tuberculosis_data_cleaned, BMI2 = wtin_kgs/((htin_cms/100)^2))
```

### Part B: Classify the BMI into `bmi_categories`

-   A healthy BMI is defined between 18,5 and 25: the person is categorized as `healthy`.

-   If the BMI is inferior to 18,5: the person is categorized as `underweight`.

-   If the BMI is between 25 and 30: the person is categorized as `overweight`.

-   If the BMI is above 30: the person is categorized as `obese`.

Using `case_when()`, create a variable, `bmi_categories` that classifies each respondent into a category. 

*Hint: Inspire yourself from the code of the conditional mutate lesson.*

```{r}
#adding a column and defining what goes into that column
tuberculosis_data_cleaned <- mutate(tuberculosis_data_cleaned,
                bmi_categories=case_when(BMI2<18.5 ~ 'underweight',
                BMI2>=18.5 & BMI2<25 ~ 'healthy',
                BMI2>=25 & BMI2<30 ~ 'overweight',
                BMI2>=30 ~ 'obese'
  ))
```

## **Step 3:** Investigating total costs

### Part A: Create a `total_cost` variable

There were three visits for the participants and each had a cost. Add together these costs to create a `total_cost` variable.

**IMPORTANT HINT:** You should `group_by()` the `id` (of participants) before using `mutate()` to sum together the different cost columns of interest.

```{r}
tuberculosis_data_cleaned <- tuberculosis_data_cleaned %>%
    group_by(id) %>%
    mutate(total_cost = sum(first_visit_cost,second_visit_cost,third_visit_cost)) %>%
ungroup()
```

### Part B: Categorize the cost 

Using `summarize()`, I will calculate for you the quantiles (0.25, 0.5 and 0.75) of the `total_cost` variable.

*For your general knowledge (although this is not a statistics course): we are using quantiles because of their capacity at splitting our data into subsets. For example, the 0.25 quantile defines the value (let's call it x) for a random variable, such that the probability that a random observation of the variable is less than x is 0.25 (25% chance).*

We will use these to define cost categories. **Run the code below**

```{r}
total_cost_quantiles <- 
  tuberculosis_data_cleaned %>%
  summarize(quantile_25 = quantile(total_cost, 0.25),
            quantile_50 = quantile(total_cost, 0.50),
            quantile_75 = quantile(total_cost, 0.75))
```

You can observe that the 0.25 quantile is at 0, the 0.5 quantile is at 500, the 0.75 quantile is at 1500. 

-   If the total cost is less than 500: the total cost is `low`.

-   If the total cost is between 500 and 1500: the total cost is `average`.

-   If the total cost is more than 1500: the total cost is `high`.

(These are arbitrary definitions based on a quick overview, that we will use for this exercise.)

Create a `total_cost_categories` variable that reflects the above classification. 

```{r}
#adding a category and defining what goes into it
tuberculosis_data_cleaned <- tuberculosis_data_cleaned %>%
  mutate(total_cost_categories = case_when(total_cost <500 ~ "low",
              total_cost >=500 & total_cost <= 1500 ~ "average",
              total_cost >1500 ~ "high"))
```

You will use this `tuberculosis_data_cleaned` for all the following subset
making and plotting.

# Tables & Plotting

## Present a demographic table

Using the sex variable and the age groups' variable you created above, print (in an aesthetic way, think `reactable()`) the average cost by age group and gender. 

*Hint: use `group_by()` and `summarize()`.*

```{r}
average_cost_gender_age_groups <- tuberculosis_data_cleaned %>%
  group_by(age_group, sex) %>%
  summarise(mean_cost = mean(total_cost))

reactable(average_cost_gender_age_groups)
```

## Cost and BMI

### Part A: Create a table

Using the gender variable, as well as the BMI category and total cost variable you created above, print (in an aesthetic way, think `reactable()`) the average cost by BMI category and gender.

```{r}
average_cost_bmi <- tuberculosis_data_cleaned %>%
  group_by(bmi_categories,sex) %>%
  summarise(mean_cost = mean(total_cost))

reactable(average_cost_bmi)
```

### Part B: Plot a histogram for the obese participants, by gender

Using the above table and `esquisse`, keep only obese respondents and plot their average total cost by gender, as a histogram. 

```{r}
average_cost_bmi %>%
 filter(bmi_categories %in% "obese" | is.na(bmi_categories)) %>%
 ggplot() +
  aes(x = sex, y = mean_cost) +
  geom_col(fill = "#112446") +
  theme_minimal()
# These data do not work as a histogram.
```

## Total costs and Treatment initiation delay

Using esquisse, plot a scatter plot of the treatment initiation delay vs. the total costs, for all participants. 

```{r}
ggplot(tuberculosis_data_cleaned) +
  aes(x = treatment_initiation_delay_in_days, y = total_cost) +
  geom_point(shape = "circle", size = 1.5, colour = "#112446") +
  theme_minimal()
```

## Employment status and Total cost categories

### Part A: Create a table

Group by employment status and the total cost categories you defined previously, to count how many participants fall in each grouping. 

```{r}
employment_cost_categories <- tuberculosis_data_cleaned %>%
  group_by(employment,total_cost_categories) %>% count()
```

### Part B: Plot it !

Select the high total cost category and compare the number of working vs non-working participants. Do so using `esquisse` and a histogram. 

Set the total cost categories on the x axis and color the histograms based on the employment status.

```{r}
ggplot(employment_cost_categories) +
  aes(x = total_cost_categories, y = n, fill = employment) +
  geom_col() +
  scale_fill_hue(direction = 1) +
  theme_minimal()

# Again, histogram doesn't work with these data
```

# Submission: Upload HTML

Once you have finished the tasks above, you should **knit this Rmd into
an HTML** and **upload it** on the assignment page.
