---
title: "Workshop 7: Pivoting then grouping"
output:
  prettydoc::html_pretty:
    theme: architect
    toc: true
    number_sections: true
author: "Zachary Anderson"
date: "2022-11-22"
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Intro

Welcome!

For this workshop, we will be cleaning a dataset. It is a hands-on
approach to using pivoting and groupings

The assignment should be submitted individually, but you are encouraged
to brainstorm with partners.

The final due date for the assignment is Tuesday, November 29th at 23:59
PM UTC+2.

# Get the assignment repo

To get started, you should download and look through the assignment
folder.

1.  First download the repo to your local computer
    [here](https://minhaskamal.github.io/DownGit/#/home?url=https://github.com/the-graph-courses/rbp_cohort_0_materials/tree/main/week_07){target="_blank"}.

    *You should ideally work on your local computer, but if you would
    rather work on RStudio Cloud, you can upload the zip file to RStudio
    Cloud through the Files pane. Consult one of the instructors for
    guidance on this.*

2.  Unzip/Extract the downloaded folder.

    *If you are on macOS, you can simply double-click on a file to unzip
    it.*

    *If you are on Windows and are not sure how to "unzip" a file, see
    [this image](https://imgur.com/a/6e5pT7k){target="_blank"}. You need
    to right-click on the file and then select "extract all".*

3.  Once done, click on the RStudio Project file in the unzipped folder
    to open the project in RStudio.

4.  In RStudio, navigate to the Files tab and open the "rmd" folder. The
    instructions for your exercise are outlined there (these are the
    same instructions you see here).

5.  Open the "data" folder and observe its components. You will work
    with the "diet_diversity_vietnam_wide_EASY.csv", "diet_diversity_vietnam_wide_INTERMEDIATE.csv" and "diet_diversity_vietnam_wide_HARD.csv" files. (The data is from the same source, remodelled for the exercise: you can also open the "00_info_about_the_dataset" file to learn more about this dataset.)


# Load packages

Now that you understand the structure of the repo, you can load in and
clean your dataset.

In the code section below, **load in the needed packages**.

```{r}
if(!require(pacman)) install.packages("pacman")
pacman::p_load(here, htmltools, reactable, tidyverse)
```

# Easy Pivoting

For this pivoting, please import the "diet_diversity_vietnam_wide_EASY.csv". The data frame you import should have 61 rows and 3 columns. Remember to use the `here()` function to allow
your Rmd to use project-relative paths.

Using the lesson on pivoting you prepared for today: pivot this dataset into long format. Print the pivoted dataframe as a reactable table.

```{r}
#Load the data
easyData <- read.csv(here("data/diet_diversity_vietnam_wide_EASY.csv"))

#Show the data before pivoting
"Before"
easyData

#Pivot the data
easyDataLong <- easyData %>%
  pivot_longer(cols = 2:3,
               names_to = "timepoint",
               values_to = "kcal"
               )

#Show the data after pivoting
"After"
reactable(easyDataLong)
```

# Intermediate Pivoting

For this pivoting, please import the "diet_diversity_vietnam_wide_INTERMEDIATE.csv". The data frame you import should have 61 rows and 5 columns. 

Using what you learnt in the code demo about some more advanced pivoting, pivot the data to long format. Print the pivoted dataframe as a reactable table.

Hint: Remember to use the neat separator in your column names.

```{r}
#Load the data
medData <- read.csv(here("data/diet_diversity_vietnam_wide_INTERMEDIATE.csv"))

#Show the data before pivoting
"Before"
medData

#Pivot the data
medDataLong <- medData %>%
  pivot_longer(!household_id,
               names_to = c("observation","timepoint"),
               names_sep = ".w_"
               )

#Show the data after pivoting
"After"
reactable(medDataLong)
```

# Bonus: Hard Pivoting (You can do it!)

For this pivoting, please import the "diet_diversity_vietnam_wide_HARD.csv". The data frame you import should have 61 rows and 9 columns. **This is the original data.**

There is no neat separator, think about how you could make one to then pivot the dataframe into long format. Print the pivoted dataframe as a reactable table.

Hint: Think about the `rename()` function of {dplyr}

```{r}
#Load the data
hardData <- read.csv(here("data/diet_diversity_vietnam_wide_HARD.csv"))


#Show the data before pivoting
"Before"
hardData

#Pivot the data
hardDataLong <- hardData %>%
  pivot_longer(!household_id,
               names_to = c("observed","timepoint"),
               names_sep = "_w_"
               )
#Despite the warning, there was an easy breakpoint in the name.
#But if there weren't, we could rename some of the columns so they all had a uniform divider, and use that divider when we pivot.

#Show the data after pivoting
"After"
reactable(hardDataLong)
```

# Submission: Upload HTML

Once you have finished the tasks above, you should **knit this Rmd into
an HTML** and **upload it** on the assignment page.
