---
title: "Honors Assignment: Extracting and Exporting Data Subsets"
output:
  prettydoc::html_pretty:
    theme: architect
    toc: true
author: "Author's name here"
date: "2022-11-01"
editor_options: 
  chunk_output_type: console
---

```{r include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Intro

Today, you will be carrying out a new task as a data analyst: preparing datasets for **someone else to use**. You can imagine this person to be a data analyst colleague or other staff member. It is a hands-on approach to using the `select()` and `filter()` verbs.

*Final due date:* **January 1st, 2023**

# Obtaining the course repo

Welcome!

Today's exercise will be done and submitted on your own, but you are encouraged to work with partners to brainstorm together.

1.  First download the course repo [here](https://minhaskamal.github.io/DownGit/#/home?url=https://github.com/the-graph-courses/rbp_cohort_0_materials/tree/main/week_04){target="_blank"}. (You will mostly work locally today).

2.  Unzip/Extract the downloaded folder. NOTE: If you are on Windows and are not sure how to "unzip" a file, see [this image](https://imgur.com/a/6e5pT7k). You need to right-click and then select "extract all".

3.  Click on the RStudio Project file to open the project in RStudio.

4.  From the Files pane of RStudio, open the "week_04" folder.

5.  In the "rmd" sub-folder of "week_04", the instructions for your exercise are outlined (these are the same instructions you see here).

6.  You will work with the "rabies_dataset" in the "data" folder. (You can view the "00_info_about_the_dataset" file to learn more about this data.)

7.  In the same folder, you also have the metadata (data dictionary) for this dataset, "rabies_metadata.pdf". Open and read through this data dictionary. Your data for today is heavily encoded (all variables are numbers corresponding to categories). Without the variable dictionary, you cannot know what the number categories correspond to!

8.  Next, perform the tasks below. Once the tasks are complete, you will zip up the "week_04" folder, which should contain (Not sure how to zip a folder? You can search Google for this.) **Your output will be several different CSV files, which you will ZIP together with this Rmd and hand in.**

------------------------------------------------------------------------

# Load and clean the data

First, **load the needed packages**. Nothing will run unless you load the packages you need to code.

Pro tip: Use {pacman} to load your packages, since it both *installs* and loads packages as needed.

```{r}
"WRITE_YOUR_CODE_HERE"
```

Now, **read in the dataset**. The data frame you import should have 1466 rows and 23 columns.

Pro tip: Try to write a path that will work on anyone's computer. (The opposite of an "absolute path").

```{r}
"WRITE_YOUR_CODE_HERE"
```

Next, **bring the respondent ID to the front of the dataset**. The respondent ID is the 23rd column of your dataset: we would like it to be more visible so that someone who opens up the CSV knows immediately that each row corresponds to a respondent.

```{r}
"WRITE_YOUR_CODE_HERE"
```

Finally, **remove the `Education` variable, which has not been properly encoded.** (To notice the issue, look both at the data frame and at the metadata file). We can consider this variable unusable for now, and remove it

```{r}
"WRITE_YOUR_CODE_HERE"
```

# Creating subsets

In each "Data subset" section below,

-   Decide whether to use the `filter()` or `select()` function, then apply it to create the required extract of the dataset.

-   Then export that dataset into an appropriately-named CSV file in the "outputs" folder.

-   NOTE: For all tasks below, in addition to the variables you should extract, please ALWAYS keep the respondent ID variable! Without it, you cannot link back your data to the original dataset and you loose crucial information.

## **Data Subset 1:** Extract demographics information

Create and export a data subset of the respondents' demographic information (age, gender and geographic background).

```{r}
"WRITE_YOUR_CODE_HERE"
```

## **Data Subset 2:** Extract all male adults from the dataset

Create and export a subset with only males aged over 18.

```{r}
"WRITE_YOUR_CODE_HERE"
```

## **Data Subset 3:** Extract at-risk individuals

Create and export a subset with "at-risk" individuals. These are people who a) have a pet at home, b) have no access to health facilities, and c) who consider that the rabies vaccine is not affordable for them.

```{r}
"WRITE_YOUR_CODE_HERE"
```

## **Data Subset 4:** Extract the knowledge-evaluation survey question variables

Reading the metadata, you will see that some variables correspond to knowledge-evaluation questions about rabies, asked to the respondents. Create and export a subset that includes these survey question variables.

```{r}
"WRITE_YOUR_CODE_HERE"
```

## **Data Subset 5:** Extract all the respondents with "ideal" knowledge, attitudes and practices (KAPs) towards rabies

People with ideal KAPs are defined as people who

-   vaccinate their pets,
-   know the signs for rabies, and
-   visit a doctor after being bitten by an animal

```{r message = F}
"WRITE_YOUR_CODE_HERE"
```
