---
title: "Fertility and Reproductive timing among Women in Nigeria"
subtitle: "Findings from the 2018 Demographic and Health Survey"
author: "GRAPH Courses team"
date: "Dec 24, 2022"
output: prettydoc::html_pretty
---

```{r echo = F}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```

```{r}
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
## Packages ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

if(!require(pacman)) install.packages("pacman")
pacman::p_load(
  janitor, # data analysis utilities
  here, # force rmds to use the project folder as working directory
  haven, # for reading in stata files
  scales, # for the comma() function, among others
  ggstatsplot,
  patchwork,
  tidyverse # for everything!
)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# ggplot2 palette  ----
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# https://mycolor.space/?hex=%2354B1BF&sub=1
my_palette <- c("#98d8e3","#74cddb","#54B1BF","#288D9B","#006B78","#004A57")

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Import and process data  ----
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

ir_raw <- haven::read_dta(here("data/NGIR7BFL.DTA"),
                          col_select = c(v012, v022, v013, v201, v213, v024, v190, 
                                         v531, v511, v212, v149, v006, v007, v009, v010, v025)
                          )




ir <- 
  ir_raw %>% 
  select(age = v012, 
         state = v022,
         age_group = v013, 
         num_kids = v201, 
         curr_pregnant = v213,
         region = v024, 
         wealth_index = v190, 
         age_first_sex = v531, 
         age_first_cohabitation = v511, 
         age_first_birth = v212, 
         educ = v149, 
         mth_interview = v006, 
         yr_interview = v007, 
         mth_birth = v009, 
         yr_birth = v010, 
         urban_rural = v025) %>% 
  haven::as_factor() %>% 
  mutate(
    educ = as.character(educ),
    educ = case_when(
      educ %in% c("incomplete primary", "complete primary") ~ "Primary",
      educ %in% c("incomplete secondary", "complete secondary") ~ "Secondary",
      educ == "no education" ~ "None",
      educ == "higher" ~ "Higher",
      TRUE ~ educ),
    educ = factor(educ, 
                  levels = c("None", "Primary", "Secondary", "Higher"))) %>% 
  mutate(wealth_index = str_to_title(wealth_index), 
         wealth_index = factor(wealth_index, 
                               levels = c("Poorest", "Poorer", "Middle", "Richer", "Richest"))) %>% 
  mutate(state = str_remove(state, " rural| urban"), 
         state = str_sub(state, start = 4), 
         state = str_to_title(state))


##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
## Fertility ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

num_women_in_survey <- nrow(ir)
num_with_child <- ir %>% filter(num_kids > 0) %>% nrow()
pct_with_child <- (100 * num_with_child/num_women_in_survey) %>% round(digits = 1)
mean_sd_kids <- 
  ir %>% summarise(mean_kids = round(mean(num_kids), digits = 2), 
                   sd_kids = round(sd(num_kids), digits = 2))

gg_col_children_born <- 
  ir %>% 
  count(num_kids) %>% 
  mutate(pct = 100 * n/sum(n),
         pct = round(pct, digits = 0), 
         pct = as.character(pct),
         pct = if_else(pct <1, "<1", pct),
         pct = paste0(pct, "%")) %>% 
  ggplot(aes(x = num_kids, y = n)) + 
  geom_col(fill = "#006B78")  +
  geom_text(aes(y = n + 400, 
                label = pct),
            size = 3.3, 
            fontface = "bold"
            #,
            #angle = 90
            )  +
  labs(x = "Children ever born", y = "Frequency count") +  
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 13500)) +
  scale_x_continuous(breaks = 0:17)+
  theme_classic()

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
## Age at first sex, marriage and birth ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

### Subset to individuals aged 15 up ----
# This is done in the 

ir_25_up <- 
  ir %>% 
  filter(age >= 25) %>% 
    mutate(age_first_sex = as.numeric(as.character(age_first_sex))) %>% 
    mutate(age_first_cohabitation = as.numeric(as.character(age_first_cohabitation))) %>% 
    mutate(age_first_birth = as.numeric(as.character(age_first_birth))) 

count_women_25_up <- ir_25_up %>% nrow() 

### Coitarche----
count_women_25_up_had_sex <- ir_25_up %>% filter(age_first_sex %in% 7:49) %>% nrow() 
pct_women_25_up_had_sex <- (100 * count_women_25_up_had_sex/count_women_25_up) %>% round(1)
median_first_sex <-
  ir_25_up %>% 
  summarise(med_age_first_sex = median(age_first_sex, na.rm = T)) %>% 
  pull(1)

### cohabiting----
count_women_25_up_have_cohabited <- ir_25_up %>% filter(!is.na(age_first_cohabitation)) %>% nrow() 
pct_women_25_up_have_cohabited <- (100 * count_women_25_up_have_cohabited/count_women_25_up) %>% round(1)
median_first_cohabitation <-
  ir_25_up %>% 
  summarise(med_age_first_cohabitation = median(age_first_cohabitation, na.rm = T)) %>% 
  pull(1)

### first birth----
count_women_25_up_have_had_child <- ir_25_up %>% filter(!is.na(age_first_birth)) %>% nrow() 
pct_women_25_up_have_had_child <- (100 * count_women_25_up_have_had_child/count_women_25_up) %>% round(1)
median_first_birth <-
  ir_25_up %>% 
  summarise(med_age_first_birth = median(age_first_birth,  na.rm = T)) %>% 
  pull(1)



### life course events plots per education and urban/rural group----

gg_col_events_per_educ_group <- 
  ir_25_up %>% 
  group_by(educ) %>% 
  summarise(med_age_first_sex = median(age_first_sex, na.rm = T), 
            med_age_first_cohabitation = median(age_first_cohabitation, na.rm = T), 
            med_age_first_birth = median(age_first_birth, na.rm = T)) %>% 
  pivot_longer(2:4) %>% 
  mutate(name = factor(name,
                       levels = c("med_age_first_sex", 
                                  "med_age_first_cohabitation", 
                                  "med_age_first_birth"))) %>% 
  ggplot() + 
  geom_col(aes(x = name, y = value, fill = educ), position = "dodge") + 
  geom_text(aes(x = name, y = value - 1.2, group = educ, label = value), 
            size = 3.8, fontface = "bold", color = "white",
            position = position_dodge(width = 0.9)) +
  scale_fill_manual(values = rev(my_palette) ) +
  scale_x_discrete(labels = c("Median age \nat first sex", 
                              "Median age \nat first cohabitation", 
                              "Median age \nat first birth")) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 27))  +
  labs(title = "Median ages at key events by education group", 
       x = "", y = "Age (years)", fill = "Education Level")  +
  theme_classic()


gg_col_events_per_urban_rural_group <- 
  ir_25_up %>% 
  group_by(urban_rural) %>% 
  summarise(med_age_first_sex = median(age_first_sex, na.rm = T), 
            med_age_first_cohabitation = median(age_first_cohabitation, na.rm = T), 
            med_age_first_birth = median(age_first_birth, na.rm = T)) %>% 
  pivot_longer(2:4) %>% 
  mutate(name = factor(name,
                       levels = c("med_age_first_sex", 
                                  "med_age_first_cohabitation", 
                                  "med_age_first_birth"))) %>% 
  mutate(urban_rural = fct_rev(urban_rural)) %>% 
  ggplot() + 
  geom_col(aes(x = name, y = value, fill = urban_rural), position = "dodge") + 
  geom_text(aes(x = name, y = value - 1.2, group = urban_rural, label = value), 
            size = 3.8, fontface = "bold", color = "white",
            position = position_dodge(width = 0.9)) +
  scale_fill_manual(values = c(my_palette[5], my_palette[3])) +
  scale_x_discrete(labels = c("Median age \nat first sex", 
                              "Median age \nat first cohabitation", 
                              "Median age \nat first birth")) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 27))  +
  labs(title = "Median ages at key events by residence type", 
       x = "", y = "Age (years)", fill = "Residence type") +
  theme_classic()

gg_events <- patchwork::wrap_plots(gg_col_events_per_educ_group, 
                      gg_col_events_per_urban_rural_group, ncol = 1) + 
  plot_annotation(tag_levels = 'a')


##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
## Teenage fertility ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
ir_teen_childbearing <- 
  ir %>% 
  filter(age_group == "15-19") %>% 
  mutate(childbearing = case_when(num_kids > 0 | curr_pregnant == "yes" ~ "Mother or pregnant",
                                  TRUE ~ "Neither mother nor pregnant"),
         childbearing = factor(childbearing, levels = c("Mother or pregnant", 
                                                        "Neither mother nor pregnant")))



# Note that this does not count towards the four ggplots required
gg_pie_childbearing  <- 
  ir_teen_childbearing %>%
  ggpiestats(
    x = childbearing,
    title = "Percent of teenagers (15-19) who have begun childbearing",
    results.subtitle = F
    ) +
  theme_void() +
  labs(fill = "Childbearing status") + 
  theme(axis.line.x.bottom = element_blank(), 
        legend.position = "right") +
  scale_fill_manual(values = c("#54B1BF", "#004A57")) + 
  guides(fill = guide_legend(reverse=T))


ir_teen_childbearing_by_region <- 
  ir_teen_childbearing %>% 
  group_by(region) %>% 
  count(childbearing) %>% 
  mutate(pct = 100 * n/sum(n)) %>% 
  filter(childbearing == "Mother or pregnant")  %>% 
  ungroup() %>% 
  mutate(region = str_to_title(region)) %>% 
  mutate(region = fct_reorder(region, pct)) %>% 
  arrange(-n)

region_hi_teen_child_rate <- ir_teen_childbearing_by_region %>% head(1)
region_lo_teen_child_rate <- ir_teen_childbearing_by_region %>% tail(1)

gg_col_ir_teen_childbearing_by_region <- 
  ir_teen_childbearing_by_region %>% 
  ggplot(aes(y = region, x = pct, fill = pct)) + 
  geom_col() + 
  scale_fill_gradient(low = "#74cddb", high = "#004A57") + 
  scale_x_continuous(expand = c(0,0)) +
  labs(title = "Teenage childbearing by region", 
       fill = "% who have \nbegun \nchildbearing", 
       x = "Percent", 
       y = "Region") +
  theme_classic() + 
  theme(axis.line.y = element_line(), 
        legend.position = "right")

gg_teen_childbearing <- 
  patchwork::wrap_plots(gg_pie_childbearing, 
                        gg_col_ir_teen_childbearing_by_region, 
                        ncol = 1) + 
  plot_annotation(tag_levels = 'a')


num_teens <- ir_teen_childbearing %>% nrow()
num_teens_childbearing <- ir_teen_childbearing %>% filter(childbearing == "Mother or pregnant") %>% nrow()
pct_teens_childbearing <- (100 * num_teens_childbearing/num_teens) %>% round(digits = 1)

```

# Fertility

Across the `r comma(num_women_in_survey)` surveyed women (ages 15 to 49), `r comma(num_with_child)`, 
(`r pct_with_child`%) had at least one child. 

```{r fig.height = 4}
#| fig.cap = "Distribution of the number of children born among surveyed women."
gg_col_children_born
```

Figure 1 shows the distribution of the number of children per women. The
mean number of children ever born was `r mean_sd_kids$mean_kids`, with a standard deviation of `r mean_sd_kids$sd_kids`.

# Age at first sex, marriage and birth

Of `r comma(count_women_25_up)` women aged 25-49 in the survey,
`r comma(count_women_25_up_had_sex)`, ( `r pct_women_25_up_had_sex`% )
have had sex, `r comma(count_women_25_up_have_cohabited)`, ( `r pct_women_25_up_had_sex`% )
have entered a cohabiting relationship (marriage or informal union), and
`r comma(count_women_25_up_have_had_child)`, ( `r pct_women_25_up_have_had_child`% ) have had at least one child.

The overall median age of first sex was `r median_first_sex` years, 
while the median first cohabitation was `r median_first_cohabitation` years, 
and the median for first birth was `r median_first_birth` years.

```{r fig.height = 8}
#| fig.cap = "Median ages at first sex, cohabitation and birth by education level and residence type. 
#| Cohabitation refers to people living together, either in a formal marraige or in an informal union"
gg_events
```

Figure 2 shows median ages at key life events by education group and residence type. 
In general, women with more education had first sex, first cohabitation and first birth later than those with less education. 
Women in urban areas also experienced these events later than those living in rural areas. 

# Teenage Fertility

```{r fig.height = 6}
#| fig.cap = "Percentage of teenage women ages 15 to 19 who have begun childbearing, overall and by region. 
#| A womam who has begun childbearing has had a child or is currently pregnant."
gg_teen_childbearing
```

Figure 3 shows the percentage of teenage women who have begun childbearing, overall and by region. 
Out of `r comma(num_teens)` teenage women (15-19) surveyed, 
`r comma(num_teens_childbearing)` (`r pct_teens_childbearing` %) are mothers or pregnant with their first child. 
By region, the teenage childbearing rate ranges from `r round(region_hi_teen_child_rate$pct)`% in `r region_hi_teen_child_rate$region` 
to  `r round(region_lo_teen_child_rate$pct)`% in `r region_lo_teen_child_rate$region`.



